---
title: "R Notebook"
output: html_notebook
---

Cluster the SH3 domain sequences, and then to compare the pattern of clustering with that of the SH3 interaction profile. 

Generate a matrix of SH3 interaction scores. 

```{r}

library(tidyverse)

# Read in the PCA data for the ABP1 swapped experiment
sh3_pca_swapped <- read_csv('PCAOct18ABP1SWAPInt_0120Analysis.csv')

# Take the data  frame given and convert the data in the form of a matrix where
# the number of rows is equal to the number of different SH3s, and the number of columns
# is equal to the number of different interactors. Therefore:

sh3_swapped <- unique(unlist(sh3_pca_swapped[,1]))
interactors_swapped <-  unique(unname(unlist(sh3_pca_swapped[,2]))) 
  
rows <- length(sh3_swapped)
cols <- length(interactors_swapped)

# Initialise matrix:
swapped_int_mat <- matrix(0,nrow=rows,ncol=cols)
rownames(swapped_int_mat) <- sh3_swapped
colnames(swapped_int_mat) <- interactors_swapped

# Assign correct matrix values

for (i in 1:length(sh3_swapped)) {
  for (j in 1:length(interactors_swapped)) {
  
    sh3_name <- sh3_swapped[i]
    int_name <- interactors_swapped[j]
  
    int_record <- sh3_pca_swapped[sh3_pca_swapped[,1] == sh3_name & sh3_pca_swapped[,2] == int_name,] 
    int_score <- unlist(unname(int_record[,5]))
    
    # If no record is found for a bait-prey pair, set interaction score to 
    # minimum across the whole dataset
    
    if(length(int_score) == 0) {int_score <- 0}
    
    swapped_int_mat[i,j] <- int_score
    
  }
}

```

Generate a distance matrix between the interaction profiles of different SH3 domains, and then perform
hierarchical clustering

```{r}

swapped_int_mat_fil <- swapped_int_mat 

# Delete 'WT ABP1' and 'ABP1 stuffed' records, which are irrelevant for this analysis 

swapped_int_mat_fil <- swapped_int_mat_fil[-which(rownames(swapped_int_mat_fil) == 'ABP1 Stuffed'),]
swapped_int_mat_fil <- swapped_int_mat_fil[-which(rownames(swapped_int_mat_fil) == 'ABP1'),]

# Generate the Euclidean distance matrix for the PPI profiles:

euclidean <- dist(swapped_int_mat_fil, method='euclidean')

```

Generate a distance matrix for the SH3 domain sequences:

```{r}

library(readxl)
library(seqinr)
detach("package:seqinr", unload=TRUE)
library(bio3d)

# Read in the SH3 domain alignment    
sh3_al <- read.fasta('bait_sh3_domain_aa_yeast_human_linsi.fasta')

#bio3d method for similarity calculation
sh3_dist_matrix <- seqidentity(sh3_al,similarity=TRUE,normalize=TRUE)
#similarities to distances
sh3_dist_matrix <- 1-sh3_dist_matrix

rownames(sh3_dist_matrix) <- toupper(rownames(sh3_dist_matrix))
colnames(sh3_dist_matrix) <- toupper(colnames(sh3_dist_matrix))

# Match row and column names between sequence matrix and PPI matrix
sh3_dist_matrix <- sh3_dist_matrix[match(rapply(strsplit(rownames(as.matrix(euclidean)),split=' '), function(x) x[1]),rownames(sh3_dist_matrix)),match(rapply(strsplit(rownames(as.matrix(euclidean)),split=' '), function(x) x[1]),rownames(sh3_dist_matrix))]

# Matrix class -> dist class                                   
sh3_dist_matrix <- as.dist(sh3_dist_matrix)

```

Directly compare the two dendrograms:

```{r}

# Use the dendextend package for this purpose

remove.packages('dendextend')
detach("package:dendextend", unload=TRUE)
devtools::install_github("DBradley27/dendextend")
library(dendextend)

# Standardise labels between the two matrices (they have to be exactly the same)
manhattan_bin <- as.matrix(euclidean)
sh3_dist_matrix <- as.matrix(sh3_dist_matrix)

rownames(manhattan_bin) <- rownames(sh3_dist_matrix)
colnames(manhattan_bin) <- colnames(sh3_dist_matrix)

# Convert matrix objects to distance objects
manhattan_bin <- as.dist(manhattan_bin)
sh3_dist_matrix <- as.dist(sh3_dist_matrix)

dend1 <- as.dendrogram(hclust(manhattan_bin))
dend2 <- as.dendrogram(hclust(sh3_dist_matrix))

# Empirical cophenetic correlation between the two matrices (PPI and SH3 domain)

empirical_score <- cor.dendlist(dendlist(dend1, dend2), method = "cophenetic")[1,2]

# Plot out the 'tangelgram'

pdf('Fig3C.pdf',9,5.5)

dendlist(dend1, dend2) %>%
  untangle(method = "step1side") %>% # Find the best alignment layout
  tanglegram(common_subtrees_color_branches = TRUE,
             highlight_branches_lwd = TRUE,
             highlight_distinct_edges = FALSE,
             common_subtrees_color_lines = TRUE,
             common_subtrees_color_lines_default_single_leaf_color = colors()[355],
             lwd = c(2.2),
             main_left = 'Interactions',
             cex_main_left = 2.1,
             main_right = 'Sequences',
             cex_main_right = 2.1,
             sort=FALSE,
             edge.lwd = 1.9,
             columns_width = c(5,3,5),
             margin_top = 4,
             margin_inner = 5,
             axes = FALSE, 
             lab.cex = 1.3,
             sub = 'Similarity clustering',
             cex_sub = 1.2)

dev.off()

```

Calculate the cophenetic correlation for permuted data in which SH3 domain labels
have been randomly assigned to each interaction profile

```{r}

# Randomisations

entanglement_vec <- NULL

for (i in 1:10000) {
  
  print(i)
  
  # Randomise assignment of different SH3 PPI profiles to different SH3 domain labels
  swapped_int_mat_fil_rand <- swapped_int_mat_fil[sample(1:nrow(swapped_int_mat_fil),nrow(swapped_int_mat_fil),replace=FALSE),]
  rownames(swapped_int_mat_fil_rand) <- rownames(swapped_int_mat_fil)
  
  # Calculate Euclidean distance matrix
  manhattan_rand <- dist(swapped_int_mat_fil_rand, method='euclidean')
  
  manhattan_rand <- as.matrix(manhattan_rand)
  sh3_dist_matrix <- as.matrix(sh3_dist_matrix)
  
  # Make row names and column names consistent between the two matrices
  rownames(manhattan_rand) <- rownames(sh3_dist_matrix)
  colnames(manhattan_rand) <- colnames(sh3_dist_matrix)
  
  # Convert matrix object to distance object
  manhattan_rand <- as.dist(manhattan_rand)
  sh3_dist_matrix <- as.dist(sh3_dist_matrix)
  
  # Generate dendrograms
  dend1 <- as.dendrogram(hclust(manhattan_rand))
  dend2 <- as.dendrogram(hclust(sh3_dist_matrix))
  
  # Calculate copehentic correlation for the randomised data
  score <- cor.dendlist(dendlist(dend1, dend2), method = "cophenetic")[1,2] 

  entanglement_vec <- c(entanglement_vec, score)
  
}

# p-value from permutation
p_value <- length(which(entanglement_vec > empirical_score))/length(entanglement_vec)

```



