---
title: "SLA_shuffle_growth_analysis"
author: "David Bradley"
date: "1/23/2020"
output:
  pdf_document: default
  html_document: default
---

1) Load in the PCA interaction data, find the Euclidean distances, and then cluster:

```{r}

library(tidyverse)

# First, read in the PCA data for the SLA1 shuffle experiments
sh3_pca_swapped <- read_csv('Sla1PPIs_270420.csv')

# Take the data frame given and convert the data in the form of a  matrix where
# the number of rows is equal to the number of different SH3s, and the number of columns
# is equal to the number of different interactors.  Therefore:

sh3_swapped <- unique(unlist(sh3_pca_swapped[,1]))
interactors_swapped <-  unique(unname(unlist(sh3_pca_swapped[,2]))) 
  
rows <- length(sh3_swapped)
cols <- length(interactors_swapped)

# Initialise matrix:
swapped_int_mat <- matrix(0,nrow=rows,ncol=cols)
rownames(swapped_int_mat) <- sh3_swapped
colnames(swapped_int_mat) <- interactors_swapped

# Assign correct matrix values

for (i in 1:length(sh3_swapped)) {
  for (j in 1:length(interactors_swapped)) {
  
    sh3_name <- sh3_swapped[i]
    int_name <- interactors_swapped[j]
  
    int_record <- sh3_pca_swapped[sh3_pca_swapped[,1] == sh3_name & sh3_pca_swapped[,2] == int_name,] 
    int_score <- unlist(unname(int_record[,5]))
    
    # If no record is found for a bait-prey pair, set interaction score to 0
    
    if(length(int_score) == 0) {int_score <- 0}

    swapped_int_mat[i,j] <- int_score
    
  }
}

```

2) Use the same approach to generate a matrix for the SLA1 growth data:

```{r}

growth_data <- read_csv('Sla1Growth_270420.csv')

# Strains

sh3_swapped_growth <- unique(unlist(growth_data[,2]))

# Remove SLA1 KO

sh3_swapped_growth <- sh3_swapped_growth[-which(sh3_swapped_growth == 'SLA1 KO')]

# Remove bad strain 'SLA1 SH3-2|SH3-3|SH3-1' from analysis:

sh3_swapped_growth <- sh3_swapped_growth[-which(sh3_swapped_growth == 'SLA1 SH3-2|SH3-3|SH3-1')]

conditions <- unique(unlist(growth_data[,1]))

rows <- length(sh3_swapped_growth)
cols <- length(conditions)

# Initialise matrix:
growth_mat <- matrix(0,nrow=rows,ncol=cols)
rownames(growth_mat) <- sh3_swapped_growth
colnames(growth_mat) <- conditions

# Assign correct matrix values

for (i in 1:length(sh3_swapped_growth)) {
  for (j in 1:length(conditions)) {
  
    sh3_name <- sh3_swapped_growth[i]
    int_name <- conditions[j]
     
    # Obtain ratio
    
    int_record <- growth_data[growth_data[,2] == sh3_name & growth_data[,1] == int_name,] 
    int_score <- unlist(unname(int_record[,5]))
    
    # If we find no record for a strain-growth combination, set score to NA
    
    if(length(int_score) == 0) {growth_mat[i,j] <- NA; next}
    
    growth_mat[i,j] <- int_score
    
  }
}

```

3) Generate a distance matrix between the interaction profiles of different SH3 domains, and then perform
hierarchical clustering. Repeat for the growth data:

```{r}

swapped_int_mat_fil <- swapped_int_mat 

# Generate the Euclidean distance matrices for the SLA1 PPI data and the SLA1 growth data:

euclidean_int <- dist(swapped_int_mat_fil, method='euclidean')

euclidean_growth <- dist(growth_mat, method='euclidean')

```

4) Directly compare the SLA1 PPI and growth dendrograms:

```{r}

remove.packages('dendextend')
detach("package:dendextend", unload=TRUE)
devtools::install_github("DBradley27/dendextend")
library(dendextend)

euclidean_int_mat <- as.matrix(euclidean_int)
euclidean_growth_mat <- as.matrix(euclidean_growth)

# Shorten row and column labels for both matrices

rownames(euclidean_growth_mat) <- rapply(strsplit(rownames(euclidean_growth_mat), split=' '), function(x) x[2])
colnames(euclidean_growth_mat) <- rapply(strsplit(colnames(euclidean_growth_mat), split=' '), function(x) x[2])
rownames(euclidean_int_mat) <- rapply(strsplit(rownames(euclidean_int_mat), split=' '), function(x) x[2])
colnames(euclidean_int_mat) <- rapply(strsplit(colnames(euclidean_int_mat), split=' '), function(x) x[2])

# Convert distance objects to matrix objects

euclidean_int_mat_dis <- as.dist(euclidean_int_mat)
euclidean_growth_mat_dis <- as.dist(euclidean_growth_mat)

# Write out the distance matrices to file

write.table(euclidean_int_mat,file='SH3_interactions_Euclidean_sla1_script_new_data.txt',quote=FALSE)
write.table(euclidean_growth_mat,file='SLA1_growth_Euclidean_new_data.txt',quote=FALSE)

euclidean_int_mat <- read.table('SH3_interactions_Euclidean_sla1_script_new_data.txt')
euclidean_growth_mat <- read.table('SLA1_growth_Euclidean_new_data.txt')

# Use default heirarchical clustering for PPI and growth data

dend1 <- as.dendrogram(hclust(as.dist(euclidean_int_mat)))
dend2 <- as.dendrogram(hclust(as.dist(euclidean_growth_mat)))

# Calculate cophenetic correlation between PPI and growth dendrograms
empirical_score <- cor.dendlist(dendlist(dend1, dend2), method = "cophenetic")[1,2]

pdf('Figs5B_new_data.pdf',9,5.5)

dendlist(dend1, dend2) %>%
  untangle(method = "step1side") %>% # Find the best alignment layout
  tanglegram(common_subtrees_color_branches = TRUE,
             highlight_branches_lwd = TRUE,
             highlight_distinct_edges = FALSE,
             common_subtrees_color_lines = TRUE,
             common_subtrees_color_lines_default_single_leaf_color = colors()[355],
             lwd = c(2.2),
             main_left = 'PPIs',
             cex_main_left = 2.2,
             main_right = 'Growth',
             cex_main_right = 2.2,
             sort=FALSE,
             edge.lwd = 1.9,
             columns_width = c(5.2,2.6,5.2),
             margin_top = 4,
             margin_inner = 9,
             axes = FALSE,
             lab.cex = 1.0,
             sub = 'Similarity clustering',
             cex_sub = 1.2)

dev.off()

```

Test the significance of the correlation by calculating the copehentic correlation for permuted data:

```{r}

entanglement_vec <- NULL

for (i in 1:100000) {
  
  print(i)
  
  # Randomise the assignment of SLA1 interaction profiles to different to different SLA1 labels
  swapped_int_mat_fil_rand <- swapped_int_mat_fil[sample(1:nrow(swapped_int_mat_fil),nrow(swapped_int_mat_fil),replace=FALSE),]
  rownames(swapped_int_mat_fil_rand) <- rownames(swapped_int_mat_fil)
  
  # Generate Euclidean distance matrix for the randomised interaction profiles
  manhattan_rand <- dist(swapped_int_mat_fil_rand, method='euclidean')
  
  # dendextend
  
  manhattan_rand <- as.matrix(manhattan_rand)

  # Convert matrix objects to distance objects
  
  manhattan_rand <- as.dist(manhattan_rand)
  euclidean_growth_mat_dis <- as.dist(euclidean_growth_mat)
  
  # Generate dendrograms for the SLA1 PPi data and growth data
  
  dend1 <- as.dendrogram(hclust(manhattan_rand))
  dend2 <- as.dendrogram(hclust(euclidean_growth_mat_dis))
  
  # Calculate a cophenetic correlation for the permuted data
  
  score <- cor.dendlist(dendlist(dend1, dend2), method = "cophenetic")[1,2] 

  entanglement_vec <- c(entanglement_vec, score)
  
}

p-value <- length(which(entanglement_vec > empirical_score))/length(entanglement_vec)

```



